Optimal (Multiway) Spatial Joins

### 2024-05-24 msj.tex
yufei

\def\vgap{\vspace{1mm}}
\def\extraspacing{\vspace{2mm} \noindent}
\def\figcapup{\vspace{-0mm}}
\def\figcapdown{\vspace{-0mm}}

@@

\def\xleft{\mathrm{left}}
\def\xright{\mathrm{right}}
\def\ybot{\mathrm{bot}}
\def\ytop{\mathrm{top}}
\def\gleft{\mathrm{left\text{-}guard}}
\def\gbot{\mathrm{bot\text{-}guard}}
\def\cross{\mathrm{cross}}
\def\out{\mathrm{OUT}}

@@

\extraspacing {\bf Sets $\bm{R_1', R_2', ..., R'_{k-2}}$.} For each rectangle $r_1 \in R_1$, we compute four segments:
\myitems{
    \item $h_1$ (resp., $h_2$): the lowest (resp., highest) segment in $H$ that crosses $r$;

    \item $v_1$ (resp., $v_2$): the leftmost (resp., rightmost) segment in $V$ that crosses $r$.
}
Define $r_1' = [x_1, x_2] \times [y_1, y_2]$, where $x_1$ (resp., $x_2$) is x-coordinate of $v_1$ (resp., $v_2$), and $y_1$ (resp., $y_2$) is y-coordinate of $h_1$ (resp., $h_2$). We say that $r_1'$ is the {\em trimmed rectangle} of $r_1$, and conversely, $r_1$ is the {\em full rectangle} of $r_1'$. Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$. Moreover, it is possible for $h_1$ and $h_2$ to be the same segment; same for $v_1$ and $v_2$. Define
\myeqn{
    R_1' &=& \set{r_1' \mid r_1 \in R_1 \text{ and its $r_1'$ defined as above exists}} \nn
}
which we call the {\em trimmed} version of $R_1$. Computing the ``segment $h_1$'' for each $r_1 \in R_1$ is an instance of Problem $\mathscr{D}$ (the find-lowest version). By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$. It thus follows from Section~\ref{sec:bricks} that $R_1'$ can be produced in $O(n \log n)$ time.

@@

In the same fashion, we obtain the trimmed version $R_i'$ of $R_i$ for all $i \in [2, k]$, spending $O(k n \log n)$ time in total. We now solve the $(k-2)$-SJ problem on $R_1', R_2', ..., R_{k-2}'$ using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}). This $(k-2)$-SJ clearly has an input size at most $n$, and let us represent its result as $\J(R_1', R_2, ..., R_{k-2})$. We prove in Appendix~\ref{app:hv}:

@@

\extraspacing {\bf Generating $\bm{\J^*}$.} Take an arbitrary $(k-2)$-tuple $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2}) \in \J(R_1', R'_2, ..., R'_{k-2})$. The reader should recall from Section~\ref{sec:bricks} that
\myitems{
    \item $B_\bm{t}$ is $\bigcap_{i=1}^{k-2} \bm{t}[i]$, i.e., the intersection of $r'_1, r'_2, ..., r'_{k-2}$;
    \item $\gleft(\bm{t})$ is the $r_i$ ($1 \le i \le k-2$) with $\xleft(r'_i) = \xleft(B_\bm{t})$;
    \item $\gbot(\bm{t})$ is the $r_i$ ($1 \le i \le k-2$) with $\ybot(r'_i) = \ybot(B_\bm{t})$.
}
We now introduce:
% \myitems{
%     \item $\cross(r_a, B; H)$, the set of segments in $H$ that cross $r_a$ and $B$;
%     \item $\cross(r_b, B; V)$, the set of segments in $V$ that cross $r_b$ and $B$.
% }
\myeqn{
    \cross(\bm{t}, H) = \set{h \in H \mid \text{$h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$}} \nn \\
    \cross(\bm{t}, V) = \set{h \in H \mid \text{$v$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}}. \nn
}
These sets have important properties as stated in the next lemma, whose proof can be found in Appendix~\ref{app:hv}:

@@

\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.}

>>>>

\def\vgap{\vspace{0mm}}
\def\extraspacing{\vspace{2mm} \noindent}
\def\figcapup{\vspace{-0mm}}
\def\figcapdown{\vspace{-0mm}}

@@

\def\xleft{\mathrm{left}}
\def\xright{\mathrm{right}}
\def\ybot{\mathrm{bot}}
\def\ytop{\mathrm{top}}
\def\maxtop{\mathrm{maxtop}}
\def\gleft{\mathrm{left\text{-}guard}}
\def\gbot{\mathrm{bot\text{-}guard}}
\def\cross{\mathrm{cross}}
\def\out{\mathrm{OUT}}

@@

\extraspacing {\bf Sets $\bm{R_1', R_2', ..., R'_{k-2}}$.} Fix any $i \in [k-2]$. For each rectangle $r \in R_i$, we compute four segments:
\myitems{
    \item $h_1$ (resp., $h_2$): the lowest (resp., highest) segment in $H$ that crosses $r$;

    \item $v_1$ (resp., $v_2$): the leftmost (resp., rightmost) segment in $V$ that crosses $r$.
}
Define $r' = [x_1, x_2] \times [y_1, y_2]$, where $x_1$ (resp., $x_2$) is x-coordinate of $v_1$ (resp., $v_2$), and $y_1$ (resp., $y_2$) is y-coordinate of $h_1$ (resp., $h_2$). We say that $r'$ is the {\em trimmed rectangle} of $r$, and conversely, $r$ is the {\em full rectangle} of $r'$. Note that $r'$ exists if and only if $r$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$. Moreover, it is possible for $h_1$ and $h_2$ to be the same segment; same for $v_1$ and $v_2$.

@@

Construct
\myeqn{
    R_i' &=& \set{r' \mid r \in R_i \text{ and its trimmed rectangle $r'$ exists}}. \nn
}
Computing the ``segment $h_1$'' for each $r \in R_1$ is an instance of Problem $\mathscr{D}$ (the find-lowest version). By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r_1 \in R_1$. It thus follows from Section~\ref{sec:bricks} that $R_i'$ can be produced in $O(n \log n)$ time.

We now solve the $(k-2)$-SJ problem on $R_1', R_2', ..., R_{k-2}'$ using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}). This $(k-2)$-SJ clearly has an input size at most $n$, and let us represent its result as $\J(R_1', R_2, ..., R_{k-2})$. We prove in Appendix~\ref{app:hv}:

@@

\extraspacing {\bf Generating $\bm{\J^*}$.} Take an arbitrary $(k-2)$-tuple $\bm{t} = (r'_1, r'_2, ...,$ $r'_{k-2}) \in \J(R_1', R'_2, ..., R'_{k-2})$. The reader should recall from Section~\ref{sec:bricks} that
\myitems{
    \item $B_\bm{t}$ is $\bigcap_{i=1}^{k-2} \bm{t}[i]$, i.e., the intersection of $r'_1, r'_2, ..., r'_{k-2}$;
    \item $\gleft(\bm{t})$ is the $r'_i$ ($1 \le i \le k-2$) with $\xleft(r'_i) = \xleft(B_\bm{t})$;
    \item $\gbot(\bm{t})$ is the $r'_i$ ($1 \le i \le k-2$) with $\ybot(r'_i) = \ybot(B_\bm{t})$.
}
We now introduce:
% \myitems{
%     \item $\cross(r_a, B; H)$, the set of segments in $H$ that cross $r_a$ and $B$;
%     \item $\cross(r_b, B; V)$, the set of segments in $V$ that cross $r_b$ and $B$.
% }
\myeqn{
    \cross(\bm{t}, H) = \set{h \in H \mid \text{$h$ crosses both $B_\bm{t}$ and $\gleft(\bm{t})$}} \nn \\
    \cross(\bm{t}, V) = \set{h \in H \mid \text{$v$ crosses both $B_\bm{t}$ and $\gbot(\bm{t})$}}. \nn
}
These sets have important properties as stated in the next lemma, whose proof can be found in Appendix~\ref{app:hv}:

@@

\extraspacing {\bf Sets $\bm{R^*_1, R^*_2, ..., R^*_{k-2}}$.} Fix any $i \in [k-2]$. For each rectangle $r' \in R_i'$, we define
\myeqn{
    \maxtop(r') &=& \max_{\bm{t} \in \J(R'_1, ..., R'_{k-2}): r' = \gbot(\bm{t})} \xright(B_\bm{t}). 
}
The value $\maxtop(r')$ is set to $-\infty$, if no $\bm{t} \in \J(R'_1, ..., R'_{k-2})$ has $r'$ as the $\gbot(\bm{t})$. When $\maxtop(r') \ne -\infty$, assuming $r' = [x_1, x_2] \times [y_1, y_2]$, we introduce a rectangle
\myeqn{
    r^* &=& [x_1, x_2] \times [y_1, \maxtop(r')]. \nn
}
We refer to $r^*$ as the {\em top-sliced rectangle} of $r'$.

Now construct from $R_i'$ a new set of rectangles: 
\myeqn{
    R^*_i = \set{r^* \mid r' \in R_i' \text{ and its top-sliced rectangle $r^*$ exists}}. \nn
}
